Radio Emission from Red-Giant Hot Jupiters

### 2015-11-07 reply2referee.txt

The nature of radio emission from red giant Hot Jupiters has already been addressed by Ignace et al (2010). The main difference in this new paper is that the authors explicitely allow for the area presented to the stellar wind by the exoplanetary magnetosphere, rather than simply the area presented by the bow shock around an unmagnetised planet. This provides around two orders of magnitude increase in the input kinetic power intercepted by the exoplanet.

This approach requires that the authors make some estimate of the magnetic field strength of the planet, for which they use published expressions for the field strength as a function of the exoplanet's mass and age. As a smaller point, in the present paper, the authors also explicitely consider the ionisation of the plasma around the exoplanet due to simple Bondi-Hoyle accretion. In the Ignace paper this mechanism was also introduced, but not explored in detail as the ionisation was provided by the presence of the bow shock that forms around the exoplanet.

The role of the size of the exoplanetary magnetosphere has been addressed by many other authors, including many of the papers by Zarka and Griessmeier cited by the authors. The novel aspect is the application to red giants. Most of the other parts of the paper involve applying published results. The most interesting of these is the use of the published expressions for the drop in the planet's convective heat flux with time as a means of estimating the decay with time of the planetary field strength and hence the emission frequency.
--------------------------------------------------------------------------------------

We thank the referee for pointing out that we were not as clear as we
would like to have been regarding the many differences from and
advances upon the seminal paper of Ignace (2010) in our original
submission.  In this resubmission, we have taken the referee's
comments to heart and have endeavored to trim literature review where
it is unnecessary, and to highlight the new aspects of our work.

### 2015-11-10 reply2referee.txt

***** TO DO LIST *****

1) response to ``The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. ''

2) argue the necesity of ``Observability`` section, in particular S4.3 and S4.4.

3) response to ``it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider. ''

**********************

@@

The nature of radio emission from red giant Hot Jupiters has already been addressed by Ignace et al (2010). The main difference in this new paper is that the authors explicitely allow for the area presented to the stellar wind by the exoplanetary magnetosphere, rather than simply the area presented by the bow shock around an unmagnetised planet. This provides around two orders of magnitude increase in the input kinetic power intercepted by the exoplanet.

This approach requires that the authors make some estimate of the magnetic field strength of the planet, for which they use published expressions for the field strength as a function of the exoplanet's mass and age. As a smaller point, in the present paper, the authors also explicitely consider the ionisation of the plasma around the exoplanet due to simple Bondi-Hoyle accretion. In the Ignace paper this mechanism was also introduced, but not explored in detail as the ionisation was provided by the presence of the bow shock that forms around the exoplanet.

The role of the size of the exoplanetary magnetosphere has been addressed by many other authors, including many of the papers by Zarka and Griessmeier cited by the authors. The novel aspect is the application to red giants. Most of the other parts of the paper involve applying published results. The most interesting of these is the use of the published expressions for the drop in the planet's convective heat flux with time as a means of estimating the decay with time of the planetary field strength and hence the emission frequency.
--------------------------------------------------------------------------------------

We thank referee's pointing this out, which made us realize that we previously failed to put our work into the context. 
As referee understands, the main topic of our paper is the re-evaluation of radio emission from planets around evolved stars. While the possibility of their strong radio emission has been addressed in Ignace et al. (2011), there are important updates in this paper, in particular:

- taking account of the shrinking of planetary magnetosphere. As a result, our estimate of radio emission is order-of-magnitude off of their estimate.  
- pointing out the presence of ionizing radition from accretion and quantitative evaluation of their ionizing effects. A mechanism to ionize otherwise neutral stellar wind has been briefly introduced in Ignace et al. (2011), but it is different from the mechanism we introduce. 
- taking account of the plasma frequency of stellar wind, which severely affects the detectability of radio signal.

In the revised manuscript, we tried to make these point clearer in Introduction section. The revised parts are marked in bold font.

Previously, we covered some topics that have already been addressed by previous papers, with an intention of making the paper self-contained and easy to follow. According to referee's suggestion, we attempted to reduce the materials as much as we think reasonable <<<grammatically wrong?>>>.  
There are some arguments which we still think are worth part of the paper. We describe them in the following, in reponse to each of referee's detailed comments.

@@

We now included there papers in this section.

--------------------------------------------------------------------------------------
Section 2.2

Eqns 6-8 seem to be missing a factor of 4pi.
--------------------------------------------------------------------------------------

We thank the referee for pointing this out. We think we missed pi for equations 7-8 instead of 4pi, since we are interested in the cross-sectional area of the stellar wind energy input. We believe eqn 6 does not need 4pi.

@@

--------------------------------------------------------------------------------------
Section 2.3

This section is only needed to calculate the decay with time of the field strength of the planet. For the radio emission, the authors simply scale to a field strength of 10G (which is perfectly acceptable as most of the parameters used in section 2.4 are unknown). This section simply quotes results from other papers and could be moved to an appendix.
--------------------------------------------------------------------------------------

The model of planetary magnetic field descibed in Section 2.3 is necessary to discuss the planetary parameter space, especially in terms of planetary mass and orbit, shown in Figure 3. Therefore, we would like to retain the description in the main manuscript. 
On the other hand, it is true that the scaling of planetary magnetic field is not a novel point of this paper (as explicitly mentioned in the manuscript), so we removed some panels from figure 2.

@@

--------------------------------------------------------------------------------------
Section 2.4

The same comment applies here...the authors ultimately only use a wind velocity that is 1/10 that of the solar wind, so much of this section is unnecessary.
--------------------------------------------------------------------------------------

It is true that some of the points discussed here are not new points. Nevertheless, we think it is our responsibility to clarify our assumptions and that readers would be rather benefited from these descriptions because these properties are not common sense in the field of exoplanets. 
It is necessary to quantitatively evaluate not only the velocity but also significant mass-loss rate. Also, the typical values of stellar wind density is helpful and necessary step to discuss plasma frequency.

@@

--------------------------------------------------------------------------------------
Section 2.5

This section is new, but not necessary, as Ignace (2010) have already addressed the issue of the ionisation of the exoplanetary exosphere. 
--------------------------------------------------------------------------------------

In Ignace et al. (2010) paper, post-shock ionization was introduced. The mechanism we newly propose in this paper provides a different view, and this allows us to adopt the same scaling law of radio emission as that of Solar-System planets. Therefore we think this is an important part of our work and we will not remove this from the paper.

@@

--------------------------------------------------------------------------------------
Section 3.2

Given that the authors dismiss the AGB stars because their cutoff frequencies are too high (which could have been demonstrated very briefly) it seems unnecessary to include these at every step. More significantly, the comparison with the "classical" hot Jupiters is not valid, either in eqns 43 or 48, as in this case the orbital velocity of the planet cannot be neglected when considering the relative velocity of the stellar wind particles to the planet (equation 7). The stellar rotational velocity is probably small and so can be neglected however. The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. 
--------------------------------------------------------------------------------------

As for the AGB stars, we believe it is important to discuss not only potentially detectable signals but also the limitation of our ability of detection. In addition, there are some parameter range where AGB companions could be detected. Therefore, we think it is worth including the discussion of AGB stars.

As for the comparison with "classical" hot Jupiters, we appreciate referee's comments. We estimated the orbital velocity of our ``canonical hot Jupiter'' at 0.05 AU to be ~130 km/sec, which can contribute 5% increase in the relative velocity. We mentioned in the paragraph following equation (43).

<<< ram pressure ??? >>>

@@

--------------------------------------------------------------------------------------
Fig 3: the numbers on the RHS have no units...they are probably fluxes, but they should be shown. 
--------------------------------------------------------------------------------------

We appreciate this feedback. We have indicated the unit in the title of Figure 3.

@@

--------------------------------------------------------------------------------------
Section 4:

Much of this section could be cut, saving perhaps section 4.2. Predictions for observations with different facilities have been made many times before.
--------------------------------------------------------------------------------------

We would like to retain Section 4.1; indeed, when we talked about this work with people, they often ask me about it. 
We also think Section 4.3 and 4.4 are useful. <<< To Tony, please describe how difficult (or not straight-forward) it is to obtain these number >>>

@@

--------------------------------------------------------------------------------------
Section 5 and Appendix B:

Could be cut.
--------------------------------------------------------------------------------------

Following the referee's comment, we cut Section 5.1, while some part of it are merged with the last section (shown in bold font). 
The intention of Section 5 was to answer a potential concern about the canceling field of far more massive stellar wind than sollar wind. We have clarified this motivation and moved it to Appendix. 
Previously we thought Appendix B might be useful for future workers (since we often got confused when reading previous papers), but we decided to cut it for the sake of brevity, as the referee suggested.

@@

--------------------------------------------------------------------------------------
Section 6:

The authors note the rareness of evolved stars - it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider.
--------------------------------------------------------------------------------------

As far as we understand the referee's feedback, Figure 5 partly answers this suggestion, where the expected flux from the known system is estimated. We now improved the discussion by including the effects of occurrence rate of Jupiter-like planets. <<< to do >>>

### 2015-11-10 ms.tex
responding to referee's comments. reply2referee.txt includes to-do list with three items

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color,xcolor}
\usepackage{url}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor={blue!50!black},citecolor={blue!50!black},urlcolor={blue!50!black}}
\allowdisplaybreaks[1]

@@

\definecolor{DarkGreen}{rgb}{0.0, 0.6, 0.0}
\def\memoTM#1{\color{DarkGreen}$[${\bf #1}$]$ \color{black}}

\def\revise#1{\color{blue}$[${\bf #1}$]$ \color{blue}}

@@

Observations of radio emission from Solar System planets imply an empirical relation that the auroral radio power is proportional to the input solar wind energy going into each planetary magnetosphere, which is commonly referred to as the ``radiometric Bode's law'' \citep{desch+kaiser1984,zarka2001}. 
Extrapolating this scaling to exoplanets, radio emission from exoplanetary systems has been examined. 
\citet{farrell1999,zarka2001,lazio2004} estimated that a few of the known exoplanetary systems may have radio flux of $\sim $1 mJy level, due to the small orbital distance of the planet and/or the larger planetary mass. 
\citet{stevens2005} gave improved estimates of the stellar mass-loss rate based on X-ray flux and re-evaluated the radio flux of known exoplanetary systems. 
\citet{griesmeier2005} took account of the high stellar activity in the early stage of the planetary system and proposed that the young system would be good candidates to search for radio emission. 
\citet{griesmeier2007a, griesmeier2007b} discussed the effects of the detailed properties of stellar wind in the proximity of the stars. They considered not only the kinetic energy of the stellar wind but also the magnetic energy of the stellar wind and coronal mass ejection (CME). 
Note that in these papers, the scaling relation of planetary magnetic fields differ from paper to paper.
In \citet{reiners2010}, the authors adopted a new scaling relation of planetary magnetic field based on \citet{christensen_et_al2009}. 
Although observational searches for these radio signatures are underway, no clear detection has been claimed \citep{bastian2000,george2007,stroe2012,hallinan2013,murphy2015}, though there are some promising initial results \citep{lecavelier_et_al2013,sirothia2014}.

@@

Based on this speculation, \citet{ignace2010} examined radio emission from known substellar-mass companions around cool evolved stars.
They found that the low ionization fraction of the stellar wind of evolved stars suppresses their radio emission and leads to weak radio emission, probably below the detection limit.

In this paper, we argue that the accretion of the massive stellar wind onto the planet would emit UV and X-ray photons that ionize the stellar wind in the vicinity of the planet.
The ionized stellar wind particles would then interact with the planetary magnetic field in the same way as Solar System planets do.
Thus, by extrapolating the ``radiometric Bode's law,'' we provide more optimistic estimates for highly evolved stars compared to the previous result.
We also consider the plasma frequency cut-off of the stellar wind, which turns out to be one of the major obstacles to detect radio emission from RGHJs. 
Given that the survey of exoplanets around highly evolved stars is not complete, we do not constrain ourselves to the systems known to have sub-stellar companions.
Instead, we investigate the parameter space where we can search for radio emission and examine the possibility to detect them with current and future instruments.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
	\plotonesc{radio_emission_mass_orbit.pdf}
   \caption{Radio flux in unit of Jy from a planetary companion to a red giant with mass loss rate $10^{-8} M_{\odot }/\mbox{yr}$ (top) and that to an AGB star with mass loss rate $10^{-5} M_{\odot }/\mbox{yr}$ (bottom). The systems are located at 100 pc away. The doubly hatched regions show the parameter spaces where the planetary radio emission would not be observable at all because the maximum frequency of the emission (cyclotron frequency at the planetary surface, $\nu_{\rm cyc}$) is below the plasma frequency cut-off, $\nu_{\rm plasma}^{\rm los}$. The regions hatched with vertical lines show the parameter spaces where the frequencies of bulk radio emission is below $\nu_{\rm plasma}^{\rm los}$, i.e., $\nu_{\rm plasma}^{\rm los} > 0.1 \nu_{\rm cyc}$)}
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We may compare radio emission power of RGHJs at 5 AU with that of canonical hot Jupiters set at 0.05 AU.
Then, equation (\ref{eq:Pkinp}) can be re-normalized as follows.
%%%
\begin{eqnarray}
P_{\rm k,\,inp} 
&\approx &  140 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_{\rm J}} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{10^{-8} M_{\odot}{\rm /yr}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for RGB stars' companions)} \notag \\
&\approx & 14000 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_{\rm J}} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{10^{-5} M_{\odot}{\rm /yr}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3}  \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for AGB stars' companions)} \notag \\
&\approx & 490 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_{\rm J}} \right)^{2/3} \left( \frac{a}{0.05~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for canonical hot Jupiters)} \notag 
\end{eqnarray}
%%%
Here we normalized the strength of magnetic field of canonical hot Jupiters with $B_J$, considering the uncertainty of the magnetic field of tidally-locked planets. 
Depending on the modeling of magnetic field strength, there is also speculation that the tidally-locked planets may have weaker magnetic fields due to slow rotation \citep[e.g.][]{griesmeier2004}; in that case the radio emission would be weaker.

@@

>>>>

\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{enumerate}
\usepackage{multirow}
\usepackage{array}
\usepackage{appendix}
\usepackage{comment}
\usepackage{color,xcolor}
\usepackage{url}
\usepackage{ulem}
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor={blue!50!black},citecolor={blue!50!black},urlcolor={blue!50!black}}
\allowdisplaybreaks[1]

@@

\definecolor{DarkGreen}{rgb}{0.0, 0.6, 0.0}
\def\memoTM#1{\color{DarkGreen}${\bf #1}$ \color{black}}

\def\revise#1#2{\color{blue}\sout{#1} {\bf #2} \color{black}}

@@

Observations of radio emission from Solar System planets imply an empirical relation that the auroral radio power is proportional to the input solar wind energy going into each planetary magnetosphere, which is commonly referred to as the ``radiometric Bode's law'' \citep{desch+kaiser1984,zarka2001}. 
Extrapolating this scaling to exoplanets, radio emission from exoplanetary systems has been examined. 
\citet{farrell1999,zarka2001,lazio2004} estimated that a few of the known exoplanetary systems may have radio flux of $\sim $1 mJy level, due to the small orbital distance of the planet and/or the larger planetary mass. 
\citet{stevens2005} gave improved estimates of the stellar mass-loss rate based on X-ray flux and re-evaluated the radio flux of known exoplanetary systems. 
\citet{griesmeier2005} took account of the high stellar activity in the early stage of the planetary system and proposed that the young system would be good candidates to search for radio emission. 
\citet{griesmeier2007a, griesmeier2007b} discussed the effects of the detailed properties of stellar wind in the proximity of the stars. They considered not only the kinetic energy of the stellar wind but also the magnetic energy of the stellar wind and coronal mass ejection (CME). 
Note that in these papers, the scaling relation of planetary magnetic fields differ from paper to paper.
In \citet{reiners2010}, the authors adopted a new scaling relation of planetary magnetic field based on \citet{christensen_et_al2009}. 
We note that \revise{}{\citet{jardine2008} modeled the reconnection between the magnetic fields of close-in planets and that of the star to obtain estimates which is not based on the ``radiometric Bode's law''. } 
Although observational searches for these radio signatures are underway, no clear detection has been claimed \revise{}{\citep{bastian2000,george2007,smith2009,stroe2012,hallinan2013,murphy2015}}, though there are some promising initial results \citep{lecavelier_et_al2013,sirothia2014}.

@@

Based on this speculation, \citet{ignace2010} examined radio emission from known substellar-mass companions around cool evolved stars.
They found that the low ionization fraction of the stellar wind of evolved stars suppresses their radio emission and leads to weak radio emission. 
\revise{}{In addition, they considered the scenario where the post-bow shock heating could produce ionized hydrogen atoms. This indicates a configuration of radio wave generation very different from that of Solar System, and the lower-limit estimate, at least, is well below the detection limit. }

In this paper, we \revise{argue that}{consider another plausible scenario where} the accretion of the massive stellar wind onto the planet would emit UV and X-ray photons that ionize the stellar wind in the vicinity of the planet.
The ionized stellar wind particles would then interact with the planetary magnetic field in the same way as Solar System planets do.
Thus, by extrapolating the ``radiometric Bode's law,'' we provide more optimistic estimates for highly evolved stars compared to the previous result. 
\revise{}{Additional two major updates from \citet{ignace2010} are the following. } 
\revise{We also consider}{Firstly, we estimate} the plasma frequency cut-off of the stellar wind, which turns out to be one of the major obstacles to detect radio emission from RGHJs. 
\revise{Given that the survey of exoplanets around highly evolved stars is not complete, we do not constrain ourselves to the systems known to have sub-stellar companions.
Instead, we investigate the parameter space where we can search for radio emission and examine the possibility to detect them with current and future instruments. }{Secondly, we discuss the planetary parameter space to  search for radio emission, employing the scaling law of planetary magnetic field, noting that the survey of exoplanets around highly evolved stars is not complete. }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[bp]
	\plotonesc{radio_emission_mass_orbit_ver2.pdf}
   \caption{Radio flux in unit of Jy from a planetary companion to a red giant with mass loss rate $10^{-8} M_{\odot }/\mbox{yr}$ (top) and that to an AGB star with mass loss rate $10^{-5} M_{\odot }/\mbox{yr}$ (bottom). The systems are located at 100 pc away. The doubly hatched regions show the parameter spaces where the planetary radio emission would not be observable at all because the maximum frequency of the emission (cyclotron frequency at the planetary surface, $\nu_{\rm cyc}$) is below the plasma frequency cut-off, $\nu_{\rm plasma}^{\rm los}$. The regions hatched with vertical lines show the parameter spaces where the frequencies of bulk radio emission is below $\nu_{\rm plasma}^{\rm los}$, i.e., $\nu_{\rm plasma}^{\rm los} > 0.1 \nu_{\rm cyc}$)}
  \label{fig:radio}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We may compare radio emission power of RGHJs at 5 AU with that of canonical hot Jupiters set at 0.05 AU.
Then, equation (\ref{eq:Pkinp}) can be re-normalized as follows.
%%%
\begin{eqnarray}
P_{\rm k,\,inp} 
&\approx &  140 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_{\rm J}} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{10^{-8} M_{\odot}{\rm /yr}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for RGB stars' companions)} \notag \\
&\approx & 14000 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_{\rm J}} \right)^{2/3} \left( \frac{a}{5~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{10^{-5} M_{\odot}{\rm /yr}} \right)^{2/3} \left( \frac{v}{10^{-1} v_{\odot}} \right)^{5/3}  \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for AGB stars' companions)} \notag \\
&\approx & 490 ~P_{\rm k,\,inp,\,J} \left( \frac{B}{B_{\rm J}} \right)^{2/3} \left( \frac{a}{0.05~\mbox{AU}} \right)^{-4/3} \notag \\
&& \times \left( \frac{\dot M_\star}{\dot M_{\odot}} \right)^{2/3} \left( \frac{v}{v_{\odot}} \right)^{5/3} \label{eq:P_HJ} \\
&& \;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \mbox{(for canonical hot Jupiters)} \notag 
\end{eqnarray}
%%%
Here we normalized the strength of magnetic field of canonical hot Jupiters with $B_J$, considering the uncertainty of the magnetic field of tidally-locked planets. 
Depending on the modeling of magnetic field strength, there is also speculation that the tidally-locked planets may have weaker magnetic fields due to slow rotation \citep[e.g.][]{griesmeier2004}; in that case the radio emission would be weaker. 
\revise{}{We also note that the orbital velocity of the canonical hot Jupiters has been ignored in equation (\ref{eq:P_HJ}); the orbital velocity at $0.05 {\rm AU}$ around a Solar-mass star is $\sim 130 {\rm km/sec}$, which could contribute to only $\sim $5\% increase of the velocity. }

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Back-reaction of a plasma flowing into a magnetic field}
\label{ss:offset}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\revise{One might expect}{Given that the density of stellar wind of evolved stars is fairly high, one might anticipate }
that \revise{}{massive} plasma flowing into a region permeated by a magnetic field would, by spiraling around the field lines, generate an opposing magnetic field that partially cancels the intrinsic planetary magnetic field. 
To estimate the strength of this effect, consider a uniform flow of particles of charge $e$, mass $m$, and velocity $v$ into a region of magnetic field $B$, spiraling along the initial magnetic field. 
A single charged particle moving perpendicular to the magnetic field moves in a circular orbit of radius 
%%%
\begin{equation}
r=\frac{mv_\bot }{eB}.
\end{equation}
%%%
which creates a magnetic dipole
%%%
\begin{equation}
\mu = \frac{e}{2\pi r/v_\bot} \pi r^2 = \frac{1}{2} e r v_\bot = \frac{mv_\bot^2}{2B}.
\end{equation}
%%%
The volume integral of the canceling magnetic field $B_c$ generated by a single dipole $\mu$ is
%%%
\begin{equation}
\int d^3{\boldsymbol r} B_c = \frac{8\pi}{3} \mu.
\end{equation}
%%%

Hence, for a number density $n$ of these charged particles, the fraction of canceling field to the background magnetic field is
%%%
\begin{equation}
\frac{B_c}{B} = \frac{n(mv_\bot^2/2)}{3 B^2/8\pi} < \frac{\rho_K}{3\rho_B},
\end{equation}
%%%
where $\rho_K$ and $\rho_B$ are the kinetic energy density and the magnetic energy density, respectively.

Just inside the stand-off point, 
%%%
\begin{equation}
m_p n[r_{\rm mag}] v^2 < \frac{B[r_{\rm mag}]^2}{2\pi }
\end{equation}
%%%
and thus $\rho_K/\rho _B < 1$. 
When the particle spirals into the planet, 
the magnetic pressures increases as 
%%%
\begin{equation}
\frac{B^2}{8\pi } \propto \frac{1}{r^6}
\end{equation}
%%%
On the other hand, the kinetic energy of the moment increases
%%%
\begin{equation}
\frac{mv^2}{2} \propto \frac{1}{r}
%\rho _K \propto \frac{1}{r}
\end{equation}
%%%
Therefore, unless the density increases more drastically than $1/r^5$, the ratio $B_c/B$ is significantly less than unity.
